---
title: "Titanic"
author: "Thiago Pires"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Introduction

This article presents some analyzes with information about the passengers of the famous ship [Titanic](https://en.wikipedia.org/wiki/RMS_Titanic) that tragically wrecked in 1912.

## Read data

```{r}
require(dplyr)
require(magrittr)
require(titanic)
train <- titanic_train
```

## About the attributes

The variables used are:

- Sex: sex
- Age: age
- Pclass: Passenger Class
- Survived: Passenger Survival Indicator

## Initial plan for data exploration

Univariate investigation to lead  after a multiple analysis. I will use plots and supervised modeling in the analysis.

## Manipulation (Actions taken for data cleaning and feature engineering)

```{r}
# tables and figures numbers
require(captioner)
fig_nums <- captioner(prefix = "Figure")
table_nums <- captioner(prefix = "Table")
# character to factor
train %<>%
  mutate(Survived = factor(x = Survived, labels = c("no", "yes")),
         Pclass = factor(x = Pclass, labels = c("1st", "2nd", "3rd")),
         Sex = factor(x = Sex))
```

## Univariate analysis

My three hypothesis about this data are:

- Relation between sex and survived
- Relation between pclass and survived
- Relation between age and survived

### Ticket class (Pclass)

There were more survivors in first class than in the second and third class (`r fig_nums("pclass", display = "cite")`).

```{r}
require(ggplot2)
train %>% count(Pclass, Survived) %>%
  group_by(Pclass) %>%
  mutate(total = sum(n)) %>%
  ggplot(aes(Pclass, n * 100/total, fill = Survived)) + geom_col() +
  theme_light() +
  labs(x = "Ticket class", y = "%", fill = "Survived",
       caption = fig_nums("pclass", "Percentual distribution of survivors according to ticket class"))
```

The model shows the odds to survive in the third class is `r round((1 - exp(-1.67039))*100, 2)` (`1 - exp(-1.67039)`) percent lower than in the first class (`r table_nums("pclass", display = "cite")`).

```{r}
require(knitr)
require(kableExtra)
model <-
  glm(Survived ~ Pclass, family = binomial, data = train)
model %>%
  summary() %$% coefficients %>%
  kable(caption = table_nums("pclass", "Model estimation"))

```

### Sex

There were more female survivors than male (`r fig_nums("sex", display = "cite")`).

```{r}
train %>% count(Sex, Survived) %>%
  group_by(Sex) %>%
  mutate(total = sum(n)) %>%
  ggplot(aes(Sex, n * 100/total, fill = Survived)) + geom_col() +
  theme_light() +
  labs(x = "Sex", y = "%", fill = "Survived",
       caption = fig_nums("sex", "Percentual distribution of survivors according to sex"))
```

The model shows the odds to survive for male is `r round((1 - exp(-2.513710))*100, 2)` (`1 - exp(-2.513710)`) percent lower than for female (`r table_nums("sex", display = "cite")`).

```{r}
model <-
  glm(Survived ~ Sex, family = binomial, data = train)
model %>%
  summary() %$% coefficients %>%
  kable(caption = table_nums("sex", "Estimativas do modelo"))

```

## Interaction between Sex and Ticket Class

The interaction between sex and ticket class was significant (`r table_nums("pclass_sex", display = "cite")`), showing that the differences between the probability of survival between men and women is greater when the ticket class improves (`r fig_nums("pclass_sex", display = "cite")`).

```{r}
model <-
  glm(Survived ~ Pclass*Sex, family = binomial, data = train)
model %>%
  summary() %$% coefficients %>%
  kable(caption = table_nums("pclass_sex", "Model estimation"))
```

```{r}
# table with all possibilities
newdata <-
  expand.grid(Pclass = c("1st", "2nd", "3rd"),
              Sex = c("male", "female")) %>% as_tibble()
# predictions (probability)
newdata %<>%
  mutate(Pihat = model %>% predict(newdata = newdata, type = "response"))
newdata %>%
  ggplot(aes(Sex, Pihat, group = Pclass, colour = Pclass)) + geom_line() + geom_point() +
  theme_light() +
  labs(x = "Sex", y = expression(pi(Survived == yes)), colour = "Ticket Class",
       caption = fig_nums("pclass_sex", "Interaction between sex and ticket class"))
```

### Age

There is no difference between the distribution of age according to survived status (`r fig_nums("age", display = "cite")`).

```{r}
train %>%
  ggplot(aes(Survived, Age)) + geom_boxplot() +
  theme_light() +
  labs(caption = fig_nums("age", "Distribution of age according to survived status"))
```

Fitting a simple model is showed a little age effect on survival probability (`r table_nums("age", display = "cite")`). The increase of one year in age decreases the odds of survival by `r round((1 - exp(-0.0109635))*100, 2)` (`1 - exp(-0.0109635)`) percent.

```{r}
model <-
  glm(Survived ~ Age, family = binomial, data = train)
model %>%
  summary() %$% coefficients %>%
  kable(caption = table_nums("age", "Model estimation"))

```

## Interaction between Age and Ticket Class

Below, note that there is no interaction between age and ticket class (p-values > .05)

```{r}
model <-
  glm(Survived ~ Pclass*Age, family = binomial, data = train)
model %>%
  summary() %$% coefficients %>%
  kable(caption = table_nums("pclass_age", "Model estimation"))
```

## Suggestions for next steps in analyzing this data

- Compare the logistic regression with another algorithms
- Chose another variables to input in the model

The variables choose there were a good quality, but age that was some missing (N = 177). To age could apply a missing imputation, this approach could be a next step to analyse too.